### 2013-03-28 paper/paper.tex
Adding LaTeX files to start writing the paper. Modify paper.tex and compile to a PDF file.

\documentclass[annual]{acmsiggraph}

\TOGonlineid{45678}
\TOGvolume{0}
\TOGnumber{0}
\TOGarticleDOI{1111111.2222222}
\TOGprojectURL{}
\TOGvideoURL{}
\TOGdataURL{}
\TOGcodeURL{}

\title{XToon}

\author{Robert Wolfe\\Spencer Elliott}
\pdfauthor{Robert Wolfe, Spencer Elliott}

\keywords{xtoon, image processing, xna}

\begin{document}

\maketitle

\begin{abstract}

Duis autem vel eum iriure dolor in hendrerit in vulputate velit esse
molestie consequat, vel illum dolore eu feugiat nulla facilisis at 
vero eros et accumsan et iusto odio dignissim qui blandit praesent
luptatum zzril delenit augue duis dolore te feugait nulla
facilisi. Lorem ipsum dolor sit amet, consectetuer adipiscing elit,
sed diam nonummy nibh euismod tincidunt ut laoreet dolore magna
aliquam erat volutpat.

Citations can be done this way~\cite{Jobs95} or this more concise 
way~\shortcite{Jobs95}, depending upon the application.

Ut wisi enim ad minim veniam, quis nostrud exerci tation ullamcorper
suscipit lobortis nisl ut aliquip ex ea commodo consequat. Duis autem
vel eum iriure dolor in hendrerit~\cite{Pellacini:2005:LAH}
in vulputate velit esse molestie~\cite{notes2002} 
consequat, vel illum dolore eu feugiat nulla facilisis at vero eros et
accumsan et iusto odio dignissim qui blandit praesent luptatum zzril
delenit augue duis dolore te feugait nulla facilisi.~\cite{Park:2006:DSI}

\end{abstract}

\keywordlist

\TOGlinkslist

\copyrightspace

\section{Introduction}

Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
minim veniam, quis nostrud exercitation ullamco laboris nisi ut
aliquip ex ea commodo consequat. Duis aute irure dolor in
reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
culpa qui officia deserunt mollit anim id est laborum.

Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
minim veniam, quis nostrud exercitation ullamco laboris nisi ut
aliquip ex ea commodo consequat. Duis aute irure dolor in
reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
culpa qui officia deserunt mollit anim id est laborum.

Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
minim veniam, quis nostrud exercitation ullamco laboris nisi ut
aliquip ex ea commodo consequat. Duis aute irure dolor in
reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
culpa qui officia deserunt mollit anim id est laborum.

\section{Tone Detail}

Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
minim veniam, quis nostrud exercitation ullamco laboris nisi ut
aliquip ex ea commodo consequat. Duis aute irure dolor in
reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
culpa qui officia deserunt mollit anim id est laborum.

Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
minim veniam, quis nostrud exercitation ullamco laboris nisi ut
aliquip ex ea commodo consequat. Duis aute irure dolor in
reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
culpa qui officia deserunt mollit anim id est laborum.

Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
minim veniam, quis nostrud exercitation ullamco laboris nisi ut
aliquip ex ea commodo consequat. Duis aute irure dolor in
reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
culpa qui officia deserunt mollit anim id est laborum.

\begin{equation}
 \sum_{j=1}^{z} j = \frac{z(z+1)}{2}
\end{equation}

\begin{eqnarray}
x & \ll & y_{1} + \cdots + y_{n} \\
  & \leq & z
\end{eqnarray}

Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
minim veniam, quis nostrud exercitation ullamco laboris nisi ut
aliquip ex ea commodo consequat. Duis aute irure dolor in
reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
culpa qui officia deserunt mollit anim id est laborum.

\section{Shape Detail}

Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
minim veniam, quis nostrud exercitation ullamco laboris nisi ut
aliquip ex ea commodo consequat. Duis aute irure dolor in
reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
culpa qui officia deserunt mollit anim id est laborum.

Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
minim veniam, quis nostrud exercitation ullamco laboris nisi ut
aliquip ex ea commodo consequat. Duis aute irure dolor in
reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
culpa qui officia deserunt mollit anim id est laborum.

Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
minim veniam, quis nostrud exercitation ullamco laboris nisi ut
aliquip ex ea commodo consequat. Duis aute irure dolor in
reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
culpa qui officia deserunt mollit anim id est laborum.

Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
minim veniam, quis nostrud exercitation ullamco laboris nisi ut
aliquip ex ea commodo consequat. Duis aute irure dolor in
reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
culpa qui officia deserunt mollit anim id est laborum.

Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
minim veniam, quis nostrud exercitation ullamco laboris nisi ut
aliquip ex ea commodo consequat. Duis aute irure dolor in
reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
culpa qui officia deserunt mollit anim id est laborum.

\section{Discussion and Future Work}

Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
minim veniam, quis nostrud exercitation ullamco laboris nisi ut
aliquip ex ea commodo consequat. Duis aute irure dolor in
reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
culpa qui officia deserunt mollit anim id est laborum.

Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
minim veniam, quis nostrud exercitation ullamco laboris nisi ut
aliquip ex ea commodo consequat. Duis aute irure dolor in
reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
culpa qui officia deserunt mollit anim id est laborum.

Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
minim veniam, quis nostrud exercitation ullamco laboris nisi ut
aliquip ex ea commodo consequat. Duis aute irure dolor in
reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
culpa qui officia deserunt mollit anim id est laborum.

Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
minim veniam, quis nostrud exercitation ullamco laboris nisi ut
aliquip ex ea commodo consequat. Duis aute irure dolor in
reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
culpa qui officia deserunt mollit anim id est laborum.

Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
minim veniam, quis nostrud exercitation ullamco laboris nisi ut
aliquip ex ea commodo consequat. Duis aute irure dolor in
reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
culpa qui officia deserunt mollit anim id est laborum.

\section{Conclusion}

Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
minim veniam, quis nostrud exercitation ullamco laboris nisi ut
aliquip ex ea commodo consequat. Duis aute irure dolor in
reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
culpa qui officia deserunt mollit anim id est laborum.

\bibliographystyle{acmsiggraph}
\bibliography{paper}
\end{document}

### 2013-03-28 comp4900-xtoon/comp4900-xtoon/Game1.cs
Added some textures and ability to switch which  texture is displayed

@@ -25,8 +25,14 @@ namespace comp4900_xtoon
         Effect celShader;
         Effect postProcessEffect;
 
+        // Models
         LinkedList<Model> models;
         LinkedListNode<Model> theModel;
+
+        // Textures
+        Texture2D Tone1DDetailTexture;
+        LinkedList<Texture2D> detailTextures;
+        LinkedListNode<Texture2D> Tone2DDetailTexture;
 
         // Used for post processing effects
         RenderTarget2D sceneRenderTarget;
@@ -41,9 +47,6 @@ namespace comp4900_xtoon
         public SpriteFont GreySpriteFont;
         public Texture2D GreyImageMap;
         public string GreyMap;
-
-        public Texture2D Tone1DDetailTexture;
-        public Texture2D Tone2DDetailTexture;
 
         public Game1()
         {
@@ -100,18 +103,25 @@ namespace comp4900_xtoon
             // Create a new SpriteBatch, which can be used to draw textures.
             spriteBatch = new SpriteBatch(GraphicsDevice);
 
+            // Models
             models = new LinkedList<Model>();
             models.AddLast(Content.Load<Model>(@"Models\dude"));
             models.AddLast(Content.Load<Model>(@"Models\ant"));
             models.AddLast(Content.Load<Model>(@"Models\Axe\FREEAXE"));
             theModel = models.First;
 
+            // Detail textures
+            Tone1DDetailTexture = Content.Load<Texture2D>(@"ToneTextures\cel_shading");
+            detailTextures = new LinkedList<Texture2D>();
+            detailTextures.AddLast(Content.Load<Texture2D>(@"ToneTextures\xtoon_white"));
+            detailTextures.AddLast(Content.Load<Texture2D>(@"ToneTextures\xtoon_shading"));
+            detailTextures.AddLast(Content.Load<Texture2D>(@"ToneTextures\xtoon_shading_alpha"));
+            Tone2DDetailTexture = detailTextures.First;
+
+            // Used for GUI
             GreyImageMap = Content.Load<Texture2D>(@"GreySkin\ImageMap");
             GreyMap = File.OpenText(@"Content\GreySkin\Map.txt").ReadToEnd();
             GreySpriteFont = Content.Load<SpriteFont>(@"GreySkin\Texture");
-
-            Tone1DDetailTexture = Content.Load<Texture2D>(@"ToneTextures\cel_shading");
-            Tone2DDetailTexture = Content.Load<Texture2D>(@"ToneTextures\xtoon_shading_alpha");
 
[…]

### 2013-03-28 comp4900-xtoon/comp4900-xtoon/Game1.cs
Added some more textures, some from the paper

@@ -114,8 +114,16 @@ namespace comp4900_xtoon
             Tone1DDetailTexture = Content.Load<Texture2D>(@"ToneTextures\cel_shading");
             detailTextures = new LinkedList<Texture2D>();
             detailTextures.AddLast(Content.Load<Texture2D>(@"ToneTextures\xtoon_white"));
-            detailTextures.AddLast(Content.Load<Texture2D>(@"ToneTextures\xtoon_shading"));
+            detailTextures.AddLast(Content.Load<Texture2D>(@"ToneTextures\xtoon_shading_gradient"));
+            detailTextures.AddLast(Content.Load<Texture2D>(@"ToneTextures\xtoon_shading_gradient2"));
             detailTextures.AddLast(Content.Load<Texture2D>(@"ToneTextures\xtoon_shading_alpha"));
+            detailTextures.AddLast(Content.Load<Texture2D>(@"ToneTextures\xtoon_shading"));
+            detailTextures.AddLast(Content.Load<Texture2D>(@"ToneTextures\xtoon_shading_grayscale"));
+            detailTextures.AddLast(Content.Load<Texture2D>(@"ToneTextures\xtoon_shading2"));
+            detailTextures.AddLast(Content.Load<Texture2D>(@"ToneTextures\xtoon_star"));
+            detailTextures.AddLast(Content.Load<Texture2D>(@"ToneTextures\xtoon_star2"));
+            detailTextures.AddLast(Content.Load<Texture2D>(@"ToneTextures\xtoon_skin"));
+            detailTextures.AddLast(Content.Load<Texture2D>(@"ToneTextures\xtoon_red"));
             Tone2DDetailTexture = detailTextures.First;
 
             // Used for GUI

### 2013-04-01 comp4900-xtoon/comp4900-xtoon/Game1.cs
Added two lights Added new GUI option to enable to use of light directions for X-Toon effect

@@ -42,6 +42,11 @@ namespace comp4900_xtoon
 
         KeyboardState oldState;
 
+        //Lighting
+        List<Light> lights;
+
+        int degrees = 0;
+
         // GUI stuff
         GuiManager gui;
         public SpriteFont GreySpriteFont;
@@ -83,6 +88,11 @@ namespace comp4900_xtoon
             // Initialize shaders
             celShader = Content.Load<Effect>(@"Effects\Cel");
             postProcessEffect = Content.Load<Effect>(@"Effects\PostProcess");
+
+            //Add 2 new lights
+            lights = new List<Light>();
+            lights.Add(new Light(new Vector3(-100.0f, 2000.0f, 100.0f)));
+            lights.Add(new Light(new Vector3(1000.0f, 2000.0f, 100.0f)));
 
             // Initialize render targets
             PresentationParameters pp = graphics.GraphicsDevice.PresentationParameters;
@@ -168,6 +178,19 @@ namespace comp4900_xtoon
             gui.Update();
             UpdateInput();
             cam.Update();
+
+            degrees++;
+            if (degrees == 360) degrees = 0;
+
+            lights[0].Position = new Vector3(
+                100.0f,
+                2000.0f + (float)Math.Cos((double)(degrees * Math.PI) / 180)*500,
+                100.0f + (float)Math.Sin((double)(degrees * Math.PI)/180)*500);
+
+            lights[1].Position = new Vector3(
+                100.0f + (float)Math.Sin((double)(degrees * Math.PI) / 180) * 500,
+                2000.0f + (float)Math.Cos((double)(degrees * Math.PI) / 180) * 500,
+                100.0f);
 
             base.Update(gameTime);
         }
@@ -278,6 +301,12 @@ namespace comp4900_xtoon
                                                         rectSize, rectSize);
                 spriteBatch.Begin();
                 spriteBatch.Draw(Tone2DDetailTexture.Value, destRectangle, Color.White);
+
+                foreach (Light l in lights)
+                {
+                    l.Draw(Tone2DDetailTexture.Value, spriteBatch);
+                }
+
                 spriteBatch.End();
             }
 
@@ -331,7 +360,18 @@ namespace comp4900_xtoon
                     effect.Parameters["Use2D"].SetValue(gui.UseXToon);
                     effect.Parameters["UseTexture"].SetValue(gui.UseTextures);
                     effect.Parameters["DetailAdjustment"].SetValue(gui.DetailAdjustment);
[…]

### 2013-04-01 comp4900-xtoon/comp4900-xtoon/GuiManager.cs

@@ -12,7 +12,7 @@ namespace comp4900_xtoon
     class GuiManager
     {
         Gui gui;
-        ToggleButton useToonButton, drawOutlineButton, useXToon, useTextureButton;
+        ToggleButton useToonButton, drawOutlineButton, useXToon, useTextureButton, useLightDirections;
         const float maxEdgeWidth = 10.0f;
         const float maxEdgeIntensity = 10.0f;
         const float maxDetail = 10000.0f;
@@ -60,7 +60,8 @@ namespace comp4900_xtoon
                     }) {
                        Value = 1 / maxDetail 
                     },
-                    detailAdjustmentLabel = new Label(margin, margin + (buttonHeight * ++i), "Detail Adjustment = 1.0")
+                    detailAdjustmentLabel = new Label(margin, margin + (buttonHeight * ++i), "Detail Adjustment = 1.0"),
+                    useLightDirections = new ToggleButton(margin, margin + (buttonHeight * ++i), "Use Light Directions")
                 }
             };
         }
@@ -98,6 +99,11 @@ namespace comp4900_xtoon
         public bool UseTextures
         {
             get { return useTextureButton.IsToggled; }
+        }
+
+        public bool UseLightDirections
+        {
+            get { return useLightDirections.IsToggled; }
         }
 
         public void Update()

### 2013-04-01 comp4900-xtoon/comp4900-xtoon/Light.cs

@@ -1,0 +1,26 @@
+﻿using System;
+using System.Collections.Generic;
+using System.Linq;
+using System.Text;
+using Microsoft.Xna.Framework;
+using Microsoft.Xna.Framework.Graphics;
+
+namespace comp4900_xtoon
+{
+    class Light
+    {
+
+        public Vector3 Position { get; set; }
+        public float Intensity { get; set; }
+
+        public Light() { }
+        public Light(Vector3 position) { Position = position; }
+        public Light(Vector3 position, float intensity) { Position = position; Intensity = intensity; }
+
+        public void Draw(Texture2D texture, SpriteBatch device)
+        {
+            Rectangle dstRect = new Rectangle((int)(Position.X), (int)(Position.Y), 20, 20);
+            device.Draw(texture, dstRect, Color.Black);
+        }
+    }
+}

### 2013-04-02 comp4900-xtoon/comp4900-xtoon/Camera.cs
Initial code to draw light positions

@@ -18,8 +18,8 @@ namespace comp4900_xtoon
         private float _cameraPitch;
         private float _aspectRatio;
 
-        private const int NEAR_CLIPPING_DISTANCE = 100;
-        private const int FAR_CLIPPING_DISTANCE = 10000;
+        public const int NEAR_CLIPPING_DISTANCE = 100;
+        public const int FAR_CLIPPING_DISTANCE = 10000;
 
         public Camera(Viewport viewport)
         {

### 2013-04-02 comp4900-xtoon/comp4900-xtoon/Game1.cs

@@ -10,6 +10,7 @@ using Microsoft.Xna.Framework.Graphics;
 using Microsoft.Xna.Framework.Input;
 using Microsoft.Xna.Framework.Media;
 using Ruminate.Utils;
+using PrimitivesSample;
 
 namespace comp4900_xtoon
 {
@@ -44,8 +45,8 @@ namespace comp4900_xtoon
 
         //Lighting
         List<Light> lights;
-
         int degrees = 0;
+        PrimitiveBatch primBatch;
 
         // GUI stuff
         GuiManager gui;
@@ -93,6 +94,8 @@ namespace comp4900_xtoon
             lights = new List<Light>();
             lights.Add(new Light(new Vector3(-100.0f, 2000.0f, 100.0f)));
             lights.Add(new Light(new Vector3(1000.0f, 2000.0f, 100.0f)));
+
+            primBatch = new PrimitiveBatch(graphics.GraphicsDevice);
 
             // Initialize render targets
             PresentationParameters pp = graphics.GraphicsDevice.PresentationParameters;
@@ -283,7 +286,14 @@ namespace comp4900_xtoon
 
             // Toon Effect
             graphics.GraphicsDevice.SetRenderTarget(sceneRenderTarget);
-            graphics.GraphicsDevice.Clear(new Color(131, 125, 151));
+            graphics.GraphicsDevice.Clear(Color.CornflowerBlue);
+
+            foreach (Light l in lights)
+            {
+                //l.Draw3D(primBatch, cam.ProjectionMatrix, new Vector3(l.Position.X, l.Position.Y, 0.0f));
+                l.Draw3D(primBatch, cam.ProjectionMatrix, new Vector3(0.0f, 0.0f, 1000.0f));
+            }
+
             DrawModel(cam.RotationMatrix, cam.ViewMatrix, cam.ProjectionMatrix, "ToonShader", theModel.Value);
 
             // Post-processing
@@ -301,12 +311,6 @@ namespace comp4900_xtoon
                                                         rectSize, rectSize);
                 spriteBatch.Begin();
                 spriteBatch.Draw(Tone2DDetailTexture.Value, destRectangle, Color.White);
-
-                foreach (Light l in lights)
-                {
-                    l.Draw(Tone2DDetailTexture.Value, spriteBatch);
-                }
-
                 spriteBatch.End();
             }
 

### 2013-04-02 comp4900-xtoon/comp4900-xtoon/Light.cs

@@ -4,11 +4,13 @@ using System.Linq;
 using System.Text;
 using Microsoft.Xna.Framework;
 using Microsoft.Xna.Framework.Graphics;
+using PrimitivesSample;
 
 namespace comp4900_xtoon
 {
     class Light
     {
+        private const int LIGHT_SIZE = 80;
 
         public Vector3 Position { get; set; }
         public float Intensity { get; set; }
@@ -21,6 +23,41 @@ namespace comp4900_xtoon
         {
             Rectangle dstRect = new Rectangle((int)(Position.X), (int)(Position.Y), 20, 20);
             device.Draw(texture, dstRect, Color.Black);
+        }
+
+        public void Draw3D(PrimitiveBatch primitiveBatch, Matrix projection, Vector3 where)
+        {
+            // the sun is made from 4 lines in a circle.
+            primitiveBatch.Begin(PrimitiveType.LineList, projection);
+
+            // draw the vertical and horizontal lines
+            primitiveBatch.AddVertex(where + new Vector3(0, LIGHT_SIZE, 0), Color.Black);
+            primitiveBatch.AddVertex(where + new Vector3(0, -LIGHT_SIZE, 0), Color.White);
+
+            primitiveBatch.AddVertex(where + new Vector3(LIGHT_SIZE, 0, 0), Color.White);
+            primitiveBatch.AddVertex(where + new Vector3(-LIGHT_SIZE, 0, 0), Color.White);
+
+            // to know where to draw the diagonal lines, we need to use trig.
+            // cosine of pi / 4 tells us what the x coordinate of a circle's radius is
+            // at 45 degrees. the y coordinate normally would come from sin, but sin and
+            // cos 45 are the same, so we can reuse cos for both x and y.
+            float sunSizeDiagonal = (float)Math.Cos(MathHelper.PiOver4);
+
+            // since that trig tells us the x and y for a unit circle, which has a
+            // radius of 1, we need scale that result by the sun's radius.
+            sunSizeDiagonal *= LIGHT_SIZE;
+
+            primitiveBatch.AddVertex(
+                where + new Vector3(-sunSizeDiagonal, sunSizeDiagonal, 0), Color.Gray);
+            primitiveBatch.AddVertex(
+                where + new Vector3(sunSizeDiagonal, -sunSizeDiagonal, 0), Color.Gray);
+
+            primitiveBatch.AddVertex(
[…]

### 2013-04-02 comp4900-xtoon/comp4900-xtoon/PrimitiveBatch.cs

@@ -1,0 +1,252 @@
+#region File Description
+//-----------------------------------------------------------------------------
+// PrimitiveBatch.cs
+//
+// Microsoft XNA Community Game Platform
+// Copyright (C) Microsoft Corporation. All rights reserved.
+//-----------------------------------------------------------------------------
+#endregion
+
+#region Using Statements
+using System;
+using System.Collections.Generic;
+using System.Text;
+using Microsoft.Xna.Framework;
+using Microsoft.Xna.Framework.Graphics;
+using Microsoft.Xna.Framework.Content;
+using Microsoft.Xna.Framework.Input;
+using comp4900_xtoon;
+#endregion
+
+namespace PrimitivesSample
+{
+
+    // PrimitiveBatch is a class that handles efficient rendering automatically for its
+    // users, in a similar way to SpriteBatch. PrimitiveBatch can render lines, points,
+    // and triangles to the screen. In this sample, it is used to draw a spacewars
+    // retro scene.
+    public class PrimitiveBatch : IDisposable
+    {
+        #region Constants and Fields
+
+        // this constant controls how large the vertices buffer is. Larger buffers will
+        // require flushing less often, which can increase performance. However, having
+        // buffer that is unnecessarily large will waste memory.
+        const int DefaultBufferSize = 500;
+
+        // a block of vertices that calling AddVertex will fill. Flush will draw using
+        // this array, and will determine how many primitives to draw from
+        // positionInBuffer.
+        VertexPositionColor[] vertices = new VertexPositionColor[DefaultBufferSize];
+
+        // keeps track of how many vertices have been added. this value increases until
+        // we run out of space in the buffer, at which time Flush is automatically
+        // called.
+        int positionInBuffer = 0;
+
+        // a basic effect, which contains the shaders that we will use to draw our
+        // primitives.
+        BasicEffect basicEffect;
+
+        // the device that we will issue draw calls to.
+        GraphicsDevice device;
+
+        // this value is set by Begin, and is the type of primitives that we are
+        // drawing.
+        PrimitiveType primitiveType;
+
+        // how many verts does each of these primitives take up? points are 1,
+        // lines are 2, and triangles are 3.
+        int numVertsPerPrimitive;
+
+        // hasBegun is flipped to true once Begin is called, and is used to make
+        // sure users don't call End before Begin is called.
+        bool hasBegun = false;
+
+        bool isDisposed = false;
+
+        #endregion
+
+        // the constructor creates a new PrimitiveBatch and sets up all of the internals
[…]

### 2013-04-04 comp4900-xtoon/comp4900-xtoon/Game1.cs
Added citations to paper and fixed camera
Made the camera easier to move. Added a test image to the paper

@@ -208,10 +208,12 @@ namespace comp4900_xtoon
             if (newState.IsKeyDown(Keys.W))
             {
                 cam.Position.X -= offset;
+                cam.LookAt.X -= offset;
             }
             if (newState.IsKeyDown(Keys.S))
             {
                 cam.Position.X += offset;
+                cam.LookAt.X += offset;
             }
             if (newState.IsKeyDown(Keys.A))
             {
@@ -225,10 +227,12 @@ namespace comp4900_xtoon
             if (newState.IsKeyDown(Keys.Down))
             {
                 cam.Position.Y -= offset;
+                cam.LookAt.Y -= offset;
             }
             if (newState.IsKeyDown(Keys.Up))
             {
                 cam.Position.Y += offset;
+                cam.LookAt.Y += offset;
             }
             if (newState.IsKeyDown(Keys.Left))
             {

### 2013-04-04 comp4900-xtoon/comp4900-xtoon/Game1.cs
Added the ability to disable lights Added light attenuation adjustment Fixed light rendering

@@ -48,6 +48,9 @@ namespace comp4900_xtoon
         int degrees = 0;
         PrimitiveBatch primBatch;
 
+        bool hideGui = false;
+        bool lastKeyUp = false;
+
         // GUI stuff
         GuiManager gui;
         public SpriteFont GreySpriteFont;
@@ -178,7 +181,7 @@ namespace comp4900_xtoon
             if (GamePad.GetState(PlayerIndex.One).Buttons.Back == ButtonState.Pressed || Keyboard.GetState(PlayerIndex.One).IsKeyDown(Keys.Escape))
                 this.Exit();
 
-            gui.Update();
+            if(!hideGui) gui.Update();
             UpdateInput();
             cam.Update();
 
@@ -187,12 +190,12 @@ namespace comp4900_xtoon
 
             lights[0].Position = new Vector3(
                 100.0f,
-                2000.0f + (float)Math.Cos((double)(degrees * Math.PI) / 180)*500,
-                100.0f + (float)Math.Sin((double)(degrees * Math.PI)/180)*500);
+                2000.0f + (float)Math.Cos((double)(degrees * Math.PI) / 180)*1000,
+                100.0f + (float)Math.Sin((double)(degrees * Math.PI)/180)*800);
 
             lights[1].Position = new Vector3(
-                100.0f + (float)Math.Sin((double)(degrees * Math.PI) / 180) * 500,
-                2000.0f + (float)Math.Cos((double)(degrees * Math.PI) / 180) * 500,
+                100.0f + (float)Math.Sin((double)((degrees + 90) * Math.PI) / 180) * 1000,
+                2000.0f + (float)Math.Cos((double)((degrees + 90) * Math.PI) / 180) * 1000,
                 100.0f);
 
             base.Update(gameTime);
@@ -247,6 +250,12 @@ namespace comp4900_xtoon
                 cam.CameraPitch += rotateAmount;
             }
 
+            //Gui hiding
+            if (oldState.IsKeyDown(Keys.H) && newState.IsKeyUp(Keys.H))
+            {
+                hideGui = !hideGui;
+            }
+
             // Models
             if (oldState.IsKeyDown(Keys.L) && newState.IsKeyUp(Keys.L))
             {
@@ -288,10 +297,13 @@ namespace comp4900_xtoon
             graphics.GraphicsDevice.SetRenderTarget(sceneRenderTarget);
             graphics.GraphicsDevice.Clear(Color.CornflowerBlue);
 
-            foreach (Light l in lights)
[…]

### 2013-04-04 comp4900-xtoon/comp4900-xtoon/GuiManager.cs

@@ -12,12 +12,13 @@ namespace comp4900_xtoon
     class GuiManager
     {
         Gui gui;
-        ToggleButton useToonButton, drawOutlineButton, useXToon, useTextureButton, useLightDirections;
+        ToggleButton useToonButton, drawOutlineButton, useXToon, useTextureButton, useLightDirections, disableLighting;
         const float maxEdgeWidth = 10.0f;
         const float maxEdgeIntensity = 10.0f;
         const float maxDetail = 10000.0f;
-        Slider edgeWidth, edgeIntensity, detailAdjustment;
-        Label edgeWidthLabel, edgeIntensityLabel, detailAdjustmentLabel;
+        const float maxAttenuation = 5000.0f;
+        Slider edgeWidth, edgeIntensity, detailAdjustment, lightAttentuation;
+        Label edgeWidthLabel, edgeIntensityLabel, detailAdjustmentLabel, lightAttenuationLabel;
 
         public void Initialize(Game1 game)
         {
@@ -61,7 +62,14 @@ namespace comp4900_xtoon
                        Value = 1 / maxDetail 
                     },
                     detailAdjustmentLabel = new Label(margin, margin + (buttonHeight * ++i), "Detail Adjustment = 1.0"),
-                    useLightDirections = new ToggleButton(margin, margin + (buttonHeight * ++i), "Use Light Directions")
+                    useLightDirections = new ToggleButton(margin, margin + (buttonHeight * ++i), "Use Light Directions"),
+                    lightAttentuation = new Slider(margin, margin + (buttonHeight * ++i), 150, delegate(Widget slider) {
+                        lightAttenuationLabel.Value = "Light attenuation = " + (((Slider)slider).Value * maxAttenuation);
+                    }) {
+                        Value = 1200 / maxAttenuation
+                    },
+                    lightAttenuationLabel = new Label(margin, margin + (buttonHeight * ++i), "Light attentuation = 1200.0"),
+                    disableLighting = new ToggleButton(margin, margin + (buttonHeight * ++i), "Disable lighting")
                 }
             };
         }
@@ -104,6 +112,16 @@ namespace comp4900_xtoon
         public bool UseLightDirections
         {
             get { return useLightDirections.IsToggled; }
+        }
+
+        public float LightAttenuation
+        {
+            get { return lightAttentuation.Value * maxAttenuation; }
+        }
+
+        public bool DisableLighting
+        {
+            get { return disableLighting.IsToggled; }
         }
 
         public void Update()

### 2013-04-04 comp4900-xtoon/comp4900-xtoon/Light.cs

@@ -25,13 +25,13 @@ namespace comp4900_xtoon
             device.Draw(texture, dstRect, Color.Black);
         }
 
-        public void Draw3D(PrimitiveBatch primitiveBatch, Matrix projection, Vector3 where)
+        public void Draw3D(PrimitiveBatch primitiveBatch, Matrix world, Matrix projection, Matrix view, Vector3 where)
         {
             // the sun is made from 4 lines in a circle.
-            primitiveBatch.Begin(PrimitiveType.LineList, projection);
+            primitiveBatch.Begin(PrimitiveType.LineList, world, projection, view);
 
             // draw the vertical and horizontal lines
-            primitiveBatch.AddVertex(where + new Vector3(0, LIGHT_SIZE, 0), Color.Black);
+            primitiveBatch.AddVertex(where + new Vector3(0, LIGHT_SIZE, 0), Color.White);
             primitiveBatch.AddVertex(where + new Vector3(0, -LIGHT_SIZE, 0), Color.White);
 
             primitiveBatch.AddVertex(where + new Vector3(LIGHT_SIZE, 0, 0), Color.White);

### 2013-04-04 paper/paper.tex

@@
 \documentclass[annual]{acmsiggraph}
-
-\TOGonlineid{45678}
-\TOGvolume{0}
-\TOGnumber{0}
-\TOGarticleDOI{1111111.2222222}
-\TOGprojectURL{}
-\TOGvideoURL{}
-\TOGdataURL{}
-\TOGcodeURL{}
 
@@
 
-\author{Robert Wolfe\\Spencer Elliott}
+\author{Robert Wolfe\thanks{e-mail:robert\_wolfe@carleton.ca}\\Spencer Elliott\thanks{e-mail:spencer\_elliott@carleton.ca}}
 \pdfauthor{Robert Wolfe, Spencer Elliott}
@@
 
-Duis autem vel eum iriure dolor in hendrerit in vulputate velit esse
-molestie consequat, vel illum dolore eu feugiat nulla facilisis at 
-vero eros et accumsan et iusto odio dignissim qui blandit praesent
-luptatum zzril delenit augue duis dolore te feugait nulla
-facilisi. Lorem ipsum dolor sit amet, consectetuer adipiscing elit,
-sed diam nonummy nibh euismod tincidunt ut laoreet dolore magna
-aliquam erat volutpat.
-
-Citations can be done this way~\cite{Jobs95} or this more concise 
-way~\shortcite{Jobs95}, depending upon the application.
-
-Ut wisi enim ad minim veniam, quis nostrud exerci tation ullamcorper
-suscipit lobortis nisl ut aliquip ex ea commodo consequat. Duis autem
-vel eum iriure dolor in hendrerit~\cite{Pellacini:2005:LAH}
-in vulputate velit esse molestie~\cite{notes2002} 
-consequat, vel illum dolore eu feugiat nulla facilisis at vero eros et
-accumsan et iusto odio dignissim qui blandit praesent luptatum zzril
-delenit augue duis dolore te feugait nulla facilisi.~\cite{Park:2006:DSI}
+We are implementing XToon ~\cite{BTM06a}.
 
@@
 
-\TOGlinkslist
-
 \copyrightspace
@@
 
-Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
-eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
-minim veniam, quis nostrud exercitation ullamco laboris nisi ut
-aliquip ex ea commodo consequat. Duis aute irure dolor in
-reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
-pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
-culpa qui officia deserunt mollit anim id est laborum.
+We first developed a standard cel shader using ~\cite{CelShadingTut}.
 
-Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
-eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
-minim veniam, quis nostrud exercitation ullamco laboris nisi ut
-aliquip ex ea commodo consequat. Duis aute irure dolor in
-reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
-pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
-culpa qui officia deserunt mollit anim id est laborum.
-
-Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
-eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
-minim veniam, quis nostrud exercitation ullamco laboris nisi ut
-aliquip ex ea commodo consequat. Duis aute irure dolor in
-reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
-pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
-culpa qui officia deserunt mollit anim id est laborum.
+We used the GUI framework for XNA from ~\cite{Ruminate}.
 
@@
 
-\section{Shape Detail}
+\section{Results}
+Here are some results
 
-Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
-eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
-minim veniam, quis nostrud exercitation ullamco laboris nisi ut
-aliquip ex ea commodo consequat. Duis aute irure dolor in
-reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
-pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
-culpa qui officia deserunt mollit anim id est laborum.
-
-Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
-eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
-minim veniam, quis nostrud exercitation ullamco laboris nisi ut
-aliquip ex ea commodo consequat. Duis aute irure dolor in
-reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
-pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
-culpa qui officia deserunt mollit anim id est laborum.
-
-Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
-eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
-minim veniam, quis nostrud exercitation ullamco laboris nisi ut
-aliquip ex ea commodo consequat. Duis aute irure dolor in
-reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
-pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
-culpa qui officia deserunt mollit anim id est laborum.
-
-Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
-eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
-minim veniam, quis nostrud exercitation ullamco laboris nisi ut
-aliquip ex ea commodo consequat. Duis aute irure dolor in
-reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
-pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
-culpa qui officia deserunt mollit anim id est laborum.
-
-Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
-eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
-minim veniam, quis nostrud exercitation ullamco laboris nisi ut
-aliquip ex ea commodo consequat. Duis aute irure dolor in
-reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
-pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
-culpa qui officia deserunt mollit anim id est laborum.
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=2.5in]{images/test}
+  \caption{Sample illustration.}
+\end{figure}
 

### 2013-04-04 comp4900-xtoon/comp4900-xtoon/Game1.cs
Adding alternative 2nd dimension for xtoon

@@ -369,6 +369,7 @@ namespace comp4900_xtoon
                     effect.Parameters["UseTexture"].SetValue(gui.UseTextures);
                     effect.Parameters["DetailAdjustment"].SetValue(gui.DetailAdjustment);
                     effect.Parameters["UseLightDirections"].SetValue(gui.UseLightDirections);
+                    effect.Parameters["LookAt"].SetValue(cam.LookAt);
 
                     //Update lights
                     int i = 0;

### 2013-04-06 comp4900-xtoon/comp4900-xtoon/Game1.cs
Worked on adding customizable lights Added example images

@@ -48,8 +48,13 @@ namespace comp4900_xtoon
         int degrees = 0;
         PrimitiveBatch primBatch;
 
+        int modelId = 0;
+        List<Vector3> lightPos = new List<Vector3>();
+        List<float> lightIntensity = new List<float>();
+        List<float> radiusSize = new List<float>();
+
         bool hideGui = false;
-        bool lastKeyUp = false;
+        bool showDetail = false;
 
         // GUI stuff
         GuiManager gui;
@@ -95,8 +100,21 @@ namespace comp4900_xtoon
 
             //Add 2 new lights
             lights = new List<Light>();
-            lights.Add(new Light(new Vector3(-100.0f, 2000.0f, 100.0f)));
-            lights.Add(new Light(new Vector3(1000.0f, 2000.0f, 100.0f)));
+            lights.Add(new Light());
+            lights.Add(new Light());
+
+            lightPos.Add(new Vector3(100.0f, 2000.0f, 100.0f));
+            lightIntensity.Add(1200.0f);
+            radiusSize.Add(1000.0f);
+
+            lightPos.Add(new Vector3(2000.0f, 2000.0f, -2000.0f));
+            lightIntensity.Add(1500.0f);
+            radiusSize.Add(1500.0f);
+
+            lightPos.Add(new Vector3(100.0f, 1000.0f, 100.0f));
+            lightIntensity.Add(800.0f);
+            radiusSize.Add(1000.0f);
+
 
             primBatch = new PrimitiveBatch(graphics.GraphicsDevice);
 
@@ -189,14 +207,17 @@ namespace comp4900_xtoon
             if (degrees == 360) degrees = 0;
 
             lights[0].Position = new Vector3(
-                100.0f,
-                2000.0f + (float)Math.Cos((double)(degrees * Math.PI) / 180)*1000,
-                100.0f + (float)Math.Sin((double)(degrees * Math.PI)/180)*800);
+                lightPos[modelId].X,
+                lightPos[modelId].Y + (float)Math.Cos((double)(degrees * Math.PI) / 180) * radiusSize[modelId],
+                lightPos[modelId].Z + (float)Math.Sin((double)(degrees * Math.PI) / 180) * radiusSize[modelId]);
 
             lights[1].Position = new Vector3(
-                100.0f + (float)Math.Sin((double)((degrees + 90) * Math.PI) / 180) * 1000,
-                2000.0f + (float)Math.Cos((double)((degrees + 90) * Math.PI) / 180) * 1000,
-                100.0f);
[…]

### 2013-04-07 comp4900-xtoon/comp4900-xtoon/Game1.cs
GUI Option to change second dimension calculation

@@ -404,6 +404,7 @@ namespace comp4900_xtoon
                     effect.Parameters["View"].SetValue(view);
                     effect.Parameters["Projection"].SetValue(projection);
                     effect.Parameters["UseToon"].SetValue(gui.UseToon);
+                    effect.Parameters["UseDistance"].SetValue(gui.UseDistance);
                     effect.Parameters["ToneTexture"].SetValue(gui.UseXToon ? Tone2DDetailTexture.Value : Tone1DDetailTexture);
                     effect.Parameters["Use2D"].SetValue(gui.UseXToon);
                     effect.Parameters["UseTexture"].SetValue(gui.UseTextures);

### 2013-04-07 comp4900-xtoon/comp4900-xtoon/Game1.cs
Added cool hologram photo Added noisy image to test

@@ -145,6 +145,7 @@ namespace comp4900_xtoon
             detailTextures.AddLast(Content.Load<Texture2D>(@"ToneTextures\xtoon_star3"));
             detailTextures.AddLast(Content.Load<Texture2D>(@"ToneTextures\xtoon_skin"));
             detailTextures.AddLast(Content.Load<Texture2D>(@"ToneTextures\xtoon_red"));
+            detailTextures.AddLast(Content.Load<Texture2D>(@"ToneTextures\noisy"));
             Tone2DDetailTexture = detailTextures.First;
 
             // Used for GUI

### 2013-04-07 paper/paper.tex

@@
 \documentclass[annual]{acmsiggraph}
-
+\usepackage{wrapfig}
 \title{XToon}
@@
 \section{Tone Detail}
+\begin{wrapfigure}{r}{0.1\textwidth}
+  \vspace{-20pt}
+  \begin{center}
+    \includegraphics[width=0.1\textwidth]{images/xtoon_skin}
+  \end{center}
+  \caption{2D Texture.}
+  \vspace{-10pt}
+  \label{fig:2dtexture}
+\end{wrapfigure}
+We extend the concept of a regular toon shader by adding a second dimension to the texture.
+While the x-axis remains as $n\cdot l$, as in a classic toon shader, the y-axis represents the level of detail.
+The higher up on the axis, the less detail is used.
+Conversely, the lower the y-axis value, the more detail.
+An example of a 2D texture used in this application can be seen in Figure~\ref{fig:2dtexture}.
 
-Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
-eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
-minim veniam, quis nostrud exercitation ullamco laboris nisi ut
-aliquip ex ea commodo consequat. Duis aute irure dolor in
-reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
-pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
-culpa qui officia deserunt mollit anim id est laborum.
+The method of computing the y-axis depends on an {\it{attribute map.}} The chosen attribute changes the level of detail, depending on its value.
+The original paper provided two possible attribute maps: distance and orientation.
+We implement distance, and probvide a new attribute map based on angle.
 
-Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
-eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
-minim veniam, quis nostrud exercitation ullamco laboris nisi ut
-aliquip ex ea commodo consequat. Duis aute irure dolor in
-reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
-pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
-culpa qui officia deserunt mollit anim id est laborum.
+\subsection{Distance}
+{\bf*{* Spencer to write in here**}}
 
-Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
-eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
-minim veniam, quis nostrud exercitation ullamco laboris nisi ut
-aliquip ex ea commodo consequat. Duis aute irure dolor in
-reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
-pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
-culpa qui officia deserunt mollit anim id est laborum.
+We use the distance along the focal axis to determine. Also light directions are used somehow.
+
+\subsection{Angle}
+We introduce a new attribute map based on the angles of the {\it{look at}} of the camera and the surface normal.
+The look at vector is the direction that the camera is pointed.
+The angle between the two vectors is calculated using Equation~\ref{eqn:angle}, which is derived from the formula for dot product.
 
 \begin{equation}
- \sum_{j=1}^{z} j = \frac{z(z+1)}{2}
+\label{eqn:angle}
+a = \cos^{-1}{(-LookAt \cdot Normal / |LookAt||Normal|)}
 \end{equation}
-
-\begin{eqnarray}
-x & \ll & y_{1} + \cdots + y_{n} \\
-  & \leq & z
-\end{eqnarray}
-
-Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
-eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
-minim veniam, quis nostrud exercitation ullamco laboris nisi ut
-aliquip ex ea commodo consequat. Duis aute irure dolor in
-reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
-pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
-culpa qui officia deserunt mollit anim id est laborum.
 

### 2013-04-10 paper/paper.tex
Worked on paper

@@
 \usepackage{wrapfig}
-\title{XToon}
+\title{X\textsuperscript{2}-Toon: An Extended X-Toon Shader}
 
@@
 
-We are implementing XToon ~\cite{BTM06a}.
+X-toon ~\cite{BTM06a} extended classic cel shading.
 
@@
 
-We first developed a standard cel shader using ~\cite{CelShadingTut}.
+We first developed a standard cel shader using ~\cite{CelShadingTut} as a basic example.
 
@@
 We extend the concept of a regular toon shader by adding a second dimension to the texture.
-While the x-axis remains as $n\cdot l$, as in a classic toon shader, the y-axis represents the level of detail.
+While the x-axis remains as $N\cdot L$, where {\it{N}} is the surface normal and {\it{L}} is the light direction, the y-axis represents the level of detail.
 The higher up on the axis, the less detail is used.
@@
 The original paper provided two possible attribute maps: distance and orientation.
-We implement distance, and probvide a new attribute map based on angle.
+This paper deals exclusively with the distance attribute mapping.
+We provide an alternative mapping involving the angle between the camera {\it{look at}} vector and surface normal.
+
+In addition to providing an alternate attribute mapping, we also developed a new method for determining the amount of shading on a model.
+This method involves calculating direction vectors to each light from each vertex.
+The resulting vector can then be used to calculate the final shading amount.
 
 \subsection{Distance}
-{\bf*{* Spencer to write in here**}}
-
-We use the distance along the focal axis to determine. Also light directions are used somehow.
+We developed a new method to determine the x-axis value based on the lighting of a scene.
+The direction to each light in the scene is determined for every vertex.
+These directions would then be summed and a resulting direction vector would be created.
+The direction vector is normalized and the final detail is calculated using the same equation for determining the
 
@@
 
+\section{Extensions to original X-Toon shader}
+We have extended the original shader offered in the X-Toon paper to increase the usability for various new effects.
+The first extension involves using colours from the original model's textures when calculating the final detail.
+The second extension allows for the use of transparency in both the tone texture and original model's textures to create
+
+\subsection{Texture blending}
+We have added the ability to blend textures from a model into the calculation of the shading and tone detail.
+This technique gives artists the ability to retain the original look of the model while also applying the X-Toon effects.
+To achieve this effect, our shader first takes the pixel sample from the original texture.
+The shading and detail are determined using the calculations explained in previous sections and the final tone detail pixel is kept.
+To mix the colours of the original texture and the tone detail pixel, a simple multiplication of the {\it{RGB}} channels is performed.
+The resulting pixel is a blend of the original texture pixel and the sampled tone detail texture.
+Figure~\ref{fig:texturing} shows an example of a model with and without texture blending.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=2.5in]{images/textures}
+	\caption{A model without and with original textures applied, respectively}
+	\label{fig:texturing}
+\end{figure}
+
+\subsection{Texture transparency}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=2.5in]{images/transparency}
+	\caption{Transparency affecting output}
+	\label{fig:transparency}
+\end{figure}
+
 \section{Results}
-Here are some results
 
@@
 
-Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
-eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
-minim veniam, quis nostrud exercitation ullamco laboris nisi ut
-aliquip ex ea commodo consequat. Duis aute irure dolor in
-reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
-pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
-culpa qui officia deserunt mollit anim id est laborum.
-
-Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
-eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
-minim veniam, quis nostrud exercitation ullamco laboris nisi ut
-aliquip ex ea commodo consequat. Duis aute irure dolor in
-reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
-pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
-culpa qui officia deserunt mollit anim id est laborum.
-
-Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
-eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
-minim veniam, quis nostrud exercitation ullamco laboris nisi ut
-aliquip ex ea commodo consequat. Duis aute irure dolor in
-reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
-pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
-culpa qui officia deserunt mollit anim id est laborum.
-
-Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
-eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
-minim veniam, quis nostrud exercitation ullamco laboris nisi ut
-aliquip ex ea commodo consequat. Duis aute irure dolor in
-reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
-pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
-culpa qui officia deserunt mollit anim id est laborum.
-
-Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
-eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
-minim veniam, quis nostrud exercitation ullamco laboris nisi ut
-aliquip ex ea commodo consequat. Duis aute irure dolor in
-reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
-pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
-culpa qui officia deserunt mollit anim id est laborum.
+\subsection{Alpha blending correction}
 
 \section{Conclusion}
-
-Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
-eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
-minim veniam, quis nostrud exercitation ullamco laboris nisi ut
-aliquip ex ea commodo consequat. Duis aute irure dolor in
-reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
-pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
-culpa qui officia deserunt mollit anim id est laborum.
 

### 2013-04-11 paper/paper.tex
Fixing compile error and fleshing out intro

@@
 \usepackage{wrapfig}
-\title{X\textsuperscript{2}-Toon: An Extended X-Toon Shader}
+\usepackage{hyperref}
+\title{X\texorpdfstring{\textsuperscript{2}}-Toon: An Extended X-Toon Shader}
 
@@
 \section{Introduction}
+~\cite{BTM06a} extended the classical understanding of how to perform toon shading.
+A new dimension of tone detail can be used to create a desired level of abstraction.
+Our approach implements this technique, as well as extends it.
+We first developed a standard cel shader using ~\cite{CelShadingTut} as a starting point.
+Using models downloaded from ~\cite{TurboSquid} and (*\textbf{where did we download the dude and axes??}), we used a threshold of the dot product of the lighting vector and the surface normal to produce classic toon shading.
+An example can be seen in Figure~\ref{fig:toonshade}.
 
-We first developed a standard cel shader using ~\cite{CelShadingTut} as a basic example.
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=1.5in]{images/classic_cel2}
+	\caption{Classic Toon Shading}
+	\label{fig:toonshade}
+\end{figure}
 
-We used the GUI framework for XNA from ~\cite{Ruminate}.
+A brief explanation of X-Toon is described in Section~\ref{sec:tonedetail}, our extensions in Section~\ref{sec:extensions}, and the results in Section~\ref{sec:results}.
+Discussion and future work are described in Section~\ref{sec:discussion} and finally the conclusion in Section~\ref{sec:conclusion}.
+
+To make our program more interactive, we developed a GUI using the XNA GUI framework from ~\cite{Ruminate}. The controls allows the user to modify the parameters and enabled/disable features.
 
 \section{Tone Detail}
+\label{sec:tonedetail}
 \begin{wrapfigure}{r}{0.1\textwidth}
@@
 \section{Extensions to original X-Toon shader}
+\label{sec:extensions}
 We have extended the original shader offered in the X-Toon paper to increase the usability for various new effects.
@@
 \section{Results}
+\label{sec:results}
 
@@
 \section{Discussion and Future Work}
+\label{sec:discussion}
 
@@
 \section{Conclusion}
+\label{sec:conclusion}
 

### 2013-04-11 paper/paper.tex
Adding images. They overlap (will fix)

@@
 Conversely, the lower the y-axis value, the more detail.
-An example of a 2D texture used in this application can be seen in Figure~\ref{fig:2dtexture}.
+An example of a 2D texture used in this application can be seen in Figure~\ref{fig:2dtexture} (note the y-axis increases downward).
 
@@
 \subsection{Distance}
-We developed a new method to determine the x-axis value based on the lighting of a scene.
-The direction to each light in the scene is determined for every vertex.
-These directions would then be summed and a resulting direction vector would be created.
-The direction vector is normalized and the final detail is calculated using the same equation for determining the
+As in the original X-Toon paper, one attribute map uses the focal distance of the vertex to the camera.
+The larger this distance, the less detail will be used, abstracting the model.
+The smaller this distance, the more detail used.
+Nothing was changed from the original paper.
+An example result can be seen in Figure~\ref{fig:distanceResult}, using a transparent texture map (see Section~\ref{sec:transparency}).
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=2.5in]{images/distance_result}
+	\caption{Parts of the model that are farther away are more transparent than closer areas.}
+	\label{fig:distanceResult}
+\end{figure}
 
@@
 \end{equation}
+
+The negative of the look at vector is used because it is pointed towards the surface (see Figure~\ref{fig:angle}).
+This value indicates what surfaces the camera is directly pointed at.
+The amount of detail used decreases as the angle between the look at vector and the surface normal approaches $180^\circ$, or when the surface is completely pointed away from the camera (or vice versa).
+As the angle approaches 0, more detail is used.
+An example result of this attribute map can be seen in Figure~\ref{fig:angleResult}.
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=2.5in]{images/angle_result}
+	\caption{An example result of the angle attribute map/}
+	\label{fig:angleResult}
+\end{figure}
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=1.5in]{images/angle}
+	\caption{Diagram of the angle calculation using the camera (left)'s look at vector and the surface normal (right). The angle difference is calculated (bottom right) using Equation~\ref{eqn:angle}.}
+	\label{fig:angle}
+\end{figure}
+
+\subsection{Light Direction}
+We developed a new method to determine the x-axis value based on the lighting of a scene.
+The direction to each light in the scene is determined for every vertex.
+These directions would then be summed and a resulting direction vector would be created.
+The direction vector is normalized and the final detail is calculated using the same equation for determining the
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=1.5in]{images/light_detail}
+	\caption{A diagram explaining the light direction attribute map.}
+	\label{fig:lightDirection}
+\end{figure}
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=3.0in]{images/light_directions}
+	\caption{An example result of using the light direction attribute map.}
+	\label{fig:lightDirectionResult}
+\end{figure}
 
@@
 
-\begin{figure}
+\begin{figure}[h]
 	\centering
-	\includegraphics[width=2.5in]{images/textures}
+	\includegraphics[width=3.0in]{images/textures}
 	\caption{A model without and with original textures applied, respectively}
@@
 \subsection{Texture transparency}
+\label{sec:transparency}
 
-\begin{figure}
+\begin{figure}[h]
 	\centering
@@
 \label{sec:results}
+Some results are ablkjalskdjalskdjasd.
 
-\begin{figure}[ht]
+\begin{figure}[h]
 	\centering
-  \includegraphics[width=2.5in]{images/test}
+	\includegraphics[width=5.5in]{images/hologram}
+	\caption{Hologram result}
+	\label{fig:hologram}
+\end{figure}
+
+\begin{figure}[p]
+	\centering
+	\includegraphics[width=5.5in]{images/inverse_lighting}
+	\caption{Inverse Lighting result}
+	\label{fig:inverseLighting}
+\end{figure}
+
+\begin{figure}[p]
+	\centering
+	\includegraphics[width=5.5in]{images/lighting}
+	\caption{Lighting result}
+	\label{fig:lightingResult}
+\end{figure}
+
+\begin{figure}[p]
+	\centering
+	\includegraphics[width=5.5in]{images/stipple-hatch-result1}
+	\caption{Stipple/hatching result}
+	\label{fig:stippleHatch1}
+\end{figure}
+
+\begin{figure}[p]
+	\centering
+	\includegraphics[width=5.5in]{images/stipple-hatch-result2}
+	\caption{Stipple/hatching result}
+	\label{fig:stippleHatch2}
+\end{figure}
+
+\begin{figure}[p]
+  \centering
+  \includegraphics[width=5.5in]{images/test}
   \caption{Sample illustration.}

### 2013-04-13 paper/paper.tex
Added explaination for transparency Started discussion section Added frame delimiting for now to get results

@@
 We extend the concept of a regular toon shader by adding a second dimension to the texture.
-While the x-axis remains as $N\cdot L$, where {\it{N}} is the surface normal and {\it{L}} is the light direction, the y-axis represents the level of detail.
+The x-axis remains as $N\cdot L$, where {\it{N}} is the surface normal and {\it{L}} is the light direction.
+The y-axis represents the level of detail.
 The higher up on the axis, the less detail is used.
@@
 \subsection{Light Direction}
-We developed a new method to determine the x-axis value based on the lighting of a scene.
-The direction to each light in the scene is determined for every vertex.
-These directions would then be summed and a resulting direction vector would be created.
-The direction vector is normalized and the final detail is calculated using the same equation for determining the
+We have developed a new method to determine the shading value on a model based on the lighting of a scene.
+Equation~\ref{eqn:light_dir} is used for determining the final vector.
+First, the direction from each light in {\it{Li}} to the vertex position, {\it{P}}, is calculated.
+These directions are then summed and a resulting direction vector is created.
+This direction is then normalized and the final detail vector, {\it{D}}, is determined.
+The same equation for determining the original shading value, $N\cdot L$, is used to determing the final shading value of the vertex.
+For the purpose of this method, {\it{N}} is represented by the final normalized direction vector, {\it{D}}.
+
+\begin{equation}
+\label{eqn:light_dir}
+D = normalize({\sum_{i=0}^{|Li|} Li[i] - P})
+\end{equation}
 
@@
 The first extension involves using colours from the original model's textures when calculating the final detail.
-The second extension allows for the use of transparency in both the tone texture and original model's textures to create
+The second extension allows for the use of transparency in both the tone texture and original model's textures to create a see-through effect for a model.
 
 \subsection{Texture blending}
+\label{sec:blending}
 We have added the ability to blend textures from a model into the calculation of the shading and tone detail.
@@
 \label{sec:transparency}
+The original X-Toon shader only supports the use of an opaque tone detail texture.
+We extend the shader to support the alpha channel of a tone detail texture to allow the use of transparency.
+Our method allows blending with original model's texture described in Section~\ref{sec:blending} or with the original method of using colour directly from the tone detail map.
+An example of transparency in a tone detail map can be seen in Figure~\ref{fig:transparency}.
+This example uses the tone detail texture shown in Figure~\ref{fig:transparent_tone}.
+The model is positioned so the detail is approximately at half.
+This gives each vertex an alpha of about 0.5 and the result is a model that is considerably transparent.
 
@@
 	\label{fig:transparency}
+\end{figure}
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=2.5in]{images/xtoon_shading_alpha}
+	\caption{Transparent tone detail texture}
+	\label{fig:transparent_tone}
 \end{figure}
@@
 
-\begin{figure}[p]
+\begin{figure}[h]
 	\centering
-	\includegraphics[width=5.5in]{images/inverse_lighting}
+	\includegraphics[width=4in]{images/inverse_lighting}
 	\caption{Inverse Lighting result}
@@
 
-\begin{figure}[p]
+\begin{figure}[h]
 	\centering
-	\includegraphics[width=5.5in]{images/lighting}
+	\includegraphics[width=4in]{images/lighting}
 	\caption{Lighting result}
@@
 
-\begin{figure}[p]
+\begin{figure}[h]
 	\centering
-	\includegraphics[width=5.5in]{images/stipple-hatch-result1}
+	\includegraphics[width=4in]{images/stipple-hatch-result1}
 	\caption{Stipple/hatching result}
@@
 
-\begin{figure}[p]
+\begin{figure}[h]
 	\centering
-	\includegraphics[width=5.5in]{images/stipple-hatch-result2}
+	\includegraphics[width=6in]{images/stipple-hatch-result2}
 	\caption{Stipple/hatching result}
@@
 
-\begin{figure}[p]
+\begin{figure}[h]
   \centering
-  \includegraphics[width=5.5in]{images/test}
+  \includegraphics[width=4in]{images/test}
   \caption{Sample illustration.}
@@
 \label{sec:discussion}
+We have presented multiple extensions to the original X-Toon shader developed by  ~\cite{BTM06a}.
+In order to retain the speed of the X-Toon shader, we attempted to minimize the complexity of the code executed in the shaders.
+The code added for transparency and texture blending as well as our new calculations for tone detail resulted in minimal performance reductions.
+To investigate the performance of our code, we used a frame counter and removed the frame limiter found in XNA.
+
+\begin{table}
+	\centering
+	\label{tab:fps}
+	\caption{Maximum frames per second SHOULD RECALCULATE}
+	\begin{tabular}{cc}
+		\hline
+    		Method  & Frames per second (max) \\
+		\hline\hline \\
+    		Original X-Toon  & 1132                       \\
+    		Light Direction & 1102                      \\
+		Angle & 1026 \\
+		Texture Blending & 1033 \\
+		Transparency & 1132 \\
+		\hline
+	\end{tabular}
+\end{table}
 
 \subsection{Alpha blending correction}
+
+\subsection{Lighting model}
 

### 2013-04-13 comp4900-xtoon/comp4900-xtoon/Game1.cs

@@ -11,6 +11,7 @@ using Microsoft.Xna.Framework.Input;
 using Microsoft.Xna.Framework.Media;
 using Ruminate.Utils;
 using PrimitivesSample;
+using System.Diagnostics;
 
 namespace comp4900_xtoon
 {
@@ -62,10 +63,23 @@ namespace comp4900_xtoon
         public Texture2D GreyImageMap;
         public string GreyMap;
 
+        Stopwatch stopWatch;
+        float maxFPS = 0.0f;
+        float avgFPS = 0.0f;
+        float lastFPS;
+        float[] fpsList = new float[100];
+        int index = 0;
+        SpriteFont font;
+
+        bool lastUseXToon = false;
+
         public Game1()
         {
             graphics = new GraphicsDeviceManager(this);
             Content.RootDirectory = "Content";
+
+            this.IsFixedTimeStep = false;
+            graphics.SynchronizeWithVerticalRetrace = false;
 
             /* In case we allow user to resize window
             Window.ClientSizeChanged += delegate
@@ -87,6 +101,9 @@ namespace comp4900_xtoon
             graphics.PreferredBackBufferHeight = 600;
             graphics.ApplyChanges();
             IsMouseVisible = true;
+
+            stopWatch = new Stopwatch();
+            stopWatch.Start();
 
             cam = new Camera(graphics.GraphicsDevice.Viewport);
 
@@ -136,6 +153,8 @@ namespace comp4900_xtoon
         {
             // Create a new SpriteBatch, which can be used to draw textures.
             spriteBatch = new SpriteBatch(GraphicsDevice);
+
+            font = Content.Load<SpriteFont>(@"timer");
 
             // Models
             models = new LinkedList<Model>();
@@ -196,6 +215,20 @@ namespace comp4900_xtoon
         /// <param name="gameTime">Provides a snapshot of timing values.</param>
         protected override void Update(GameTime gameTime)
         {
+            if (lastUseXToon != gui.UseXToon)
+            {
+                lastUseXToon = gui.UseXToon;
+
+                fpsList = new float[100];
+                index = 0;
+                lastFPS = 0;
+                avgFPS = 0;
+                maxFPS = 0;
+            }
+
+            avgFPS = CalcAvgFPS(lastFPS);
+            if (lastFPS > maxFPS && !float.IsInfinity(lastFPS)) maxFPS = lastFPS;
+
             // Allows the game to exit
[…]

### 2013-04-14 paper/paper.tex
Finished sections of paper (needs some editing) Added fullscreen code (commented out) Put time step back in

@@
 
-X-toon ~\cite{BTM06a} extended classic cel shading.
+The X-toon shader ~\cite{BTM06a} extended classic cel shading from 1D shading texture to 2D texture.
+The second dimension "corresponds to the desired 'tone detail'".
+In this paper, we look to extend the initial offering of the X-Toon shader and develop new methods for calculating shading values and tone detail.
+Additionally, we extend the original shader to allow the use of model textures and transparency within a tone detail texture.
+This gives artists the tools to create new effects that were not possible with original X-Toon shader.
 
@@
 We first developed a standard cel shader using ~\cite{CelShadingTut} as a starting point.
-Using models downloaded from ~\cite{TurboSquid} and (*\textbf{where did we download the dude and axes??}), we used a threshold of the dot product of the lighting vector and the surface normal to produce classic toon shading.
+Using models downloaded from ~\cite{TurboSquid}, we used a threshold of the dot product of the lighting vector and the surface normal to produce classic toon shading.
 An example can be seen in Figure~\ref{fig:toonshade}.
@@
 \end{wrapfigure}
-We extend the concept of a regular toon shader by adding a second dimension to the texture.
+X-Toon extends the regular toon shader by adding a second dimension to the texture.
 The x-axis remains as $N\cdot L$, where {\it{N}} is the surface normal and {\it{L}} is the light direction.
@@
 \subsection{Distance}
+\label{sec:distance}
 As in the original X-Toon paper, one attribute map uses the focal distance of the vertex to the camera.
@@
 \subsection{Angle}
+\label{sec:angle}
 We introduce a new attribute map based on the angles of the {\it{look at}} of the camera and the surface normal.
@@
 \subsection{Light Direction}
+\label{sec:light_dir}
 We have developed a new method to determine the shading value on a model based on the lighting of a scene.
@@
 \label{sec:results}
-Some results are ablkjalskdjalskdjasd.
+Using the new methods outlined in Sections~\ref{sec:angle}, ~\ref{sec:light_dir}, ~\ref{sec:blending}, and ~\ref{sec:transparency}, we were able to develop new effects not possible with the original X-Toon shader.
+Figures~\ref{fig:hologram}, ~\ref{fig:stippleHatch1}, and ~\ref{fig:stippleHatch2} use the same tone detail texture but result in different effects.
+This is achieved by mixing our techniques such as using the lighting directions outlined in ~\ref{sec:light_dir} to create a hologram effect in Figure~\ref{fig:hologram}.
+Figure~\ref{fig:lightingResult} simulates self-shadowing.
+We use the textures from the original models and blend them with the tone detail texture found in Figure~\ref{fig:gradient}.
+This uses our texture blending technique described in Section~\ref{sec:blending}.
+Additionally, we were also able to have the self-shadowing effect interact with the lighting of the scene using the lighting directions as previously discussed.
 
@@
 	\centering
+	\includegraphics[width=4in]{images/xtoon_shading_gradient}
+	\caption{Tone texture for lighting}
+	\label{fig:gradient}
+\end{figure}
+
+\begin{figure}[h]
+	\centering
 	\includegraphics[width=4in]{images/stipple-hatch-result1}
@@
   \includegraphics[width=4in]{images/test}
-  \caption{Sample illustration.}
+  \caption{Sample illustration}
+\end{figure}
+
+\begin{figure}[h]
+ \centering
+ \includegraphics[width=4in]{images/alpha_issues}
+ \caption{Issues with alpha blending. The vertices on the front of the model can still be seen.}
+ \label{fig:alpha_issue}
 \end{figure}
@@
 We have presented multiple extensions to the original X-Toon shader developed by  ~\cite{BTM06a}.
-In order to retain the speed of the X-Toon shader, we attempted to minimize the complexity of the code executed in the shaders.
-The code added for transparency and texture blending as well as our new calculations for tone detail resulted in minimal performance reductions.
-To investigate the performance of our code, we used a frame counter and removed the frame limiter found in XNA.
+In order to retain the speed of the X-Toon shader, we attempted to minimize the complexity of the code executed in our shaders.
+The code added for transparency and texture blending, as well as our new calculations for tone detail, resulted in minimal performance reductions.
+To investigate the performance of our code, we calculated the frames per second (FPS) and removed the fixed time step.
+Two nVidia Geforce GTX 470 in SLI mode and an Intel i5 processor on Windows 8 were used for the calculation.
+The values found in Table~\ref{tab:fps} show the resulting FPS of the new methods introduced in this paper compared to the original X-Toon shader.
+We have concluded that our results are within a suitable error threshold and do not hinder the original shader, performance-wise.
+This data does not take into account any errors that may have arisen from factors such as the operating system and background applications affecting the outcome.
+In the two next sections, we will discuss future work that can be done to improve upon our methods outlined in this paper.
+First, we describe issues with alpha blending and a possible solution.
+Second, we discuss the effect of adding multiple lights to a scene and their performance implications.
 
@@
 	\label{tab:fps}
-	\caption{Maximum frames per second SHOULD RECALCULATE}
-	\begin{tabular}{cc}
+	\caption{Frames per second (FPS) of each new technique compared to original X-Toon shading}
+	\begin{tabular}{cccc}
 		\hline
-    		Method  & Frames per second (max) \\
+    		Method  & FPS (average) & FPS (max) & \% difference \\
 		\hline\hline \\
-    		Original X-Toon  & 1132                       \\
-    		Light Direction & 1102                      \\
-		Angle & 1026 \\
-		Texture Blending & 1033 \\
-		Transparency & 1132 \\
+    		Original X-Toon  & 1348 & 1420  & -                     \\
+    		Light Direction & 1323 & 1435 & -1.85                     \\
+		Angle & 1354 & 1447 & 0.45 \\
+		Texture Blending & 1347 & 1430 & -0.07 \\
+		Transparency & 1345 & 1419 & -0.22 \\
 		\hline
@@
 \subsection{Alpha blending correction}
+Figure~\ref{fig:alpha_issue} demonstrates issues with alpha blending when using a tone detail texture with transparency.
+When a model is rendered, the polygons hidden from the camera's view should not be rendered.
+Unfortunately, we are using alpha blending while rendering which results in these polygons being visible.
+In future works, a different technique should be used to calculate whether a pixel should overwrite the original pixel colour based on the depth and position of the polygon.
+If a polygon is fully covered and has a deeper depth value, it should be excluded from the scene.
+The alpha channel should only be considered in the output pixel value returned from the pixel shader.
 
 \subsection{Lighting model}
+The most significant result pertains to the light direction technique outlined in Section~\ref{sec:light_dir}.
+The calculated difference from Table~\ref{tab:fps} resulted in a 1.85\% performance drop compared to the original X-Toon shader.
+This is the result of only using two lights for the detail calculation.
+This performance drop will only increase as more lights are added to a scene.
+When calculating the light direction, the shader loops over each light in the scene resulting in a linear complexity of $O(n)$.
+In future works, a different lighting model should be developed to interact with the shader.
+The method we have developed is not viable for complex scenes and is limited to a very small amount of lights before it begins to reduce performance significantly.
 
@@
 \label{sec:conclusion}
+We have presented multiple extensions to the original X-Toon shader outlined by ~\cite{BTM06a}.
+Our texture blending method allow for artists to blend the original textures from a model with the calculated tone detail.
+Additionally, we have extended the original tone detail textures to allow transparency which creates see-through effects for a model.
+We have also outlined a new method for determining the shading based on lighting direction.
+Finally, a new method was created to calculate detail based on the the angle between a camera's look at vector and a model's surface normal.
+Combining the techniques outlined in this paper, a designer can achieve many new effects not possible in the original X-Toon shader.
 

### 2013-04-14 comp4900-xtoon/comp4900-xtoon/Game1.cs

@@ -20,6 +20,8 @@ namespace comp4900_xtoon
     /// </summary>
     public class Game1 : Microsoft.Xna.Framework.Game
     {
+        const int MAX_SAMPLES = 500;
+
         GraphicsDeviceManager graphics;
         SpriteBatch spriteBatch;
 
@@ -66,8 +68,9 @@ namespace comp4900_xtoon
         Stopwatch stopWatch;
         float maxFPS = 0.0f;
         float avgFPS = 0.0f;
+        long curSum = 0;
         float lastFPS;
-        float[] fpsList = new float[100];
+        float[] fpsList = new float[MAX_SAMPLES];
         int index = 0;
         SpriteFont font;
 
@@ -78,8 +81,13 @@ namespace comp4900_xtoon
             graphics = new GraphicsDeviceManager(this);
             Content.RootDirectory = "Content";
 
-            this.IsFixedTimeStep = false;
-            graphics.SynchronizeWithVerticalRetrace = false;
+            //this.IsFixedTimeStep = false;
+            //graphics.SynchronizeWithVerticalRetrace = false;
+
+            //graphics.PreferredBackBufferWidth = 1680;
+            //graphics.PreferredBackBufferHeight = 1050;
+            //graphics.PreferMultiSampling = false;
+            //graphics.IsFullScreen = true;
 
             /* In case we allow user to resize window
             Window.ClientSizeChanged += delegate
@@ -219,11 +227,12 @@ namespace comp4900_xtoon
             {
                 lastUseXToon = gui.UseXToon;
 
-                fpsList = new float[100];
+                fpsList = new float[MAX_SAMPLES];
                 index = 0;
                 lastFPS = 0;
                 avgFPS = 0;
                 maxFPS = 0;
+                curSum = 0;
             }
 
             avgFPS = CalcAvgFPS(lastFPS);
@@ -348,18 +357,18 @@ namespace comp4900_xtoon
 
         public float CalcAvgFPS(float newFPS)
         {
+            curSum -= (long)fpsList[index];
             fpsList[index] = newFPS;
             index++;
 
-            if (index >= 100) index = 0;
-
-            float sum = 0.0f;
-            foreach (float f in fpsList)
+            if (index >= MAX_SAMPLES)
             {
-                sum += f;
+                index = 0;
             }
+            
+            curSum += (long)newFPS;
 
-            return sum / 100;
+            return (float)(curSum / MAX_SAMPLES);
         }
 
         /// <summary>

### 2013-04-15 paper/paper.tex
Fixed some issues in paper

@@
 The X-toon shader ~\cite{BTM06a} extended classic cel shading from 1D shading texture to 2D texture.
-The second dimension "corresponds to the desired 'tone detail'".
+The second dimension ``corresponds to the desired `tone detail' ''.
 In this paper, we look to extend the initial offering of the X-Toon shader and develop new methods for calculating shading values and tone detail.
@@
 
-To make our program more interactive, we developed a GUI using the XNA GUI framework from ~\cite{Ruminate}. The controls allows the user to modify the parameters and enabled/disable features.
+To make our program more interactive, we developed a GUI using the XNA GUI framework from ~\cite{Ruminate}. The controls allows the user to modify the parameters and enable/disable features.
 
@@
 \label{sec:angle}
-We introduce a new attribute map based on the angles of the {\it{look at}} of the camera and the surface normal.
+We introduce a new attribute map based on the angles of the {\it{look at}} vector of the camera and the surface normal.
 The look at vector is the direction that the camera is pointed.
@@
 \label{sec:results}
-Using the new methods outlined in Sections~\ref{sec:angle}, ~\ref{sec:light_dir}, ~\ref{sec:blending}, and ~\ref{sec:transparency}, we were able to develop new effects not possible with the original X-Toon shader.
-Figures~\ref{fig:hologram}, ~\ref{fig:stippleHatch1}, and ~\ref{fig:stippleHatch2} use the same tone detail texture but result in different effects.
-This is achieved by mixing our techniques such as using the lighting directions outlined in ~\ref{sec:light_dir} to create a hologram effect in Figure~\ref{fig:hologram}.
+Using the new methods outlined in Sections~\ref{sec:angle},~\ref{sec:light_dir},~\ref{sec:blending}, and~\ref{sec:transparency}, we were able to develop new effects not possible with the original X-Toon shader.
+Figures~\ref{fig:hologram},~\ref{fig:stippleHatch1}, and~\ref{fig:stippleHatch2} use the same tone detail texture but result in different effects.
+This is achieved by mixing our techniques such as using the lighting directions outlined in~\ref{sec:light_dir} to create a hologram effect in Figure~\ref{fig:hologram}.
 Figure~\ref{fig:lightingResult} simulates self-shadowing.
@@
 	\centering
+	\caption{Frames per second (FPS) of each new technique compared to original X-Toon shading}
 	\label{tab:fps}
-	\caption{Frames per second (FPS) of each new technique compared to original X-Toon shading}
 	\begin{tabular}{cccc}

### 2013-04-15 paper/paper.tex
Fixed images so they take up the whole page

@@
 	\centering
-	\includegraphics[width=2.5in]{images/xtoon_shading_alpha}
-	\caption{Transparent tone detail texture}
+	\includegraphics[width=1.5in]{images/xtoon_shading_alpha}
+	\caption{Transparent tone detail texture. Note: the gray represents the transparency in the image.}
 	\label{fig:transparent_tone}
@@
 
-\begin{figure}[h]
+\begin{figure*}[h]
 	\centering
@@
 	\label{fig:hologram}
-\end{figure}
+\end{figure*}
+
+\begin{figure*}[h]
+	\centering
+	\includegraphics[width=5.5in]{images/inverse_lighting}
+	\caption{Inverse Lighting result}
+	\label{fig:inverseLighting}
+\end{figure*}
+
+\begin{figure*}[h]
+	\centering
+	\includegraphics[width=5.5in]{images/lighting}
+	\caption{Lighting result}
+	\label{fig:lightingResult}
+\end{figure*}
 
@@
 	\centering
-	\includegraphics[width=4in]{images/inverse_lighting}
-	\caption{Inverse Lighting result}
-	\label{fig:inverseLighting}
-\end{figure}
-
-\begin{figure}[h]
-	\centering
-	\includegraphics[width=4in]{images/lighting}
-	\caption{Lighting result}
-	\label{fig:lightingResult}
-\end{figure}
-
-\begin{figure}[h]
-	\centering
-	\includegraphics[width=4in]{images/xtoon_shading_gradient}
+	\includegraphics[width=1.5in]{images/xtoon_shading_gradient}
 	\caption{Tone texture for lighting}
@@
 
-\begin{figure}[h]
+\begin{figure*}[h]
 	\centering
-	\includegraphics[width=4in]{images/stipple-hatch-result1}
+	\includegraphics[width=5.5in]{images/stipple-hatch-result1}
 	\caption{Stipple/hatching result}
 	\label{fig:stippleHatch1}
-\end{figure}
+\end{figure*}
 
-\begin{figure}[h]
+\begin{figure*}[h]
 	\centering
-	\includegraphics[width=6in]{images/stipple-hatch-result2}
+	\includegraphics[width=5.5in]{images/stipple-hatch-result2}
 	\caption{Stipple/hatching result}
 	\label{fig:stippleHatch2}
-\end{figure}
+\end{figure*}
 
-\begin{figure}[h]
+\begin{figure*}[h]
   \centering
-  \includegraphics[width=4in]{images/test}
+  \includegraphics[width=5.5in]{images/test}
   \caption{Sample illustration}
-\end{figure}
+\end{figure*}
 
-\begin{figure}[h]
+\begin{figure*}[h]
  \centering
- \includegraphics[width=4in]{images/alpha_issues}
+ \includegraphics[width=5.5in]{images/alpha_issues}
  \caption{Issues with alpha blending. The vertices on the front of the model can still be seen.}
  \label{fig:alpha_issue}
-\end{figure}
+\end{figure*}
 
@@
 		\hline
-    		Method  & FPS (average) & FPS (max) & \% difference \\
+    		Method  & FPS (avg) & FPS (max) & \% difference \\
 		\hline\hline \\

### 2013-04-15 paper/paper.tex

@@
 	\includegraphics[width=5.5in]{images/hologram}
-	\caption{Hologram result}
+	\caption{A hologram-like result. Textures from the model are not used, and a noise texture is used. As the y-axis on the 2D Texture decreases, the amount of noise increases.}
 	\label{fig:hologram}
@@
 	\includegraphics[width=5.5in]{images/stipple-hatch-result1}
-	\caption{Stipple/hatching result}
+	\caption{A stipple and hatching-like result of an ant.
+A noise 2D texture was used.
+Vertices closer to the camera were more noisy, while those further away were transparent.
+The gaster of the ant looks like it was done through hatching, while the rest of its body looks like it was done through stipppling.}
 	\label{fig:stippleHatch1}
@@
 	\includegraphics[width=5.5in]{images/stipple-hatch-result2}
-	\caption{Stipple/hatching result}
+	\caption{A hatching-like result of an ant. A noise 2D texture was used. Vertices closer to the camera were more noisy, and looked more like hatches, while those further away were transparent.}
 	\label{fig:stippleHatch2}
@@
   \includegraphics[width=5.5in]{images/test}
-  \caption{Sample illustration}
+  \caption{Using a 2D texture from the original paper, we create a unique result by mixing the model's texture colors and the colors from the detail map.}
 \end{figure*}

### 2013-04-15 paper/paper.tex
More fixes to the paper

@@
 	\centering
-	\includegraphics[width=1.5in]{images/classic_cel2}
+	\includegraphics[width=2.5in]{images/classic_cel2}
 	\caption{Classic Toon Shading}
@@
 Conversely, the lower the y-axis value, the more detail.
-An example of a 2D texture used in this application can be seen in Figure~\ref{fig:2dtexture} (note the y-axis increases downward).
+An example of a 2D tone detail texture used in this application can be seen in Figure~\ref{fig:2dtexture} (note the y-axis increases downward).
 
@@
 Nothing was changed from the original paper.
-An example result can be seen in Figure~\ref{fig:distanceResult}, using a transparent texture map (see Section~\ref{sec:transparency}).
+An example result can be seen in Figure~\ref{fig:distanceResult}, using a transparent tone detail texture (see Section~\ref{sec:transparency}).
 
@@
 As the angle approaches 0, more detail is used.
-An example result of this attribute map can be seen in Figure~\ref{fig:angleResult}.
+An example result of this attribute map using a transparent tone detail texture can be seen in Figure~\ref{fig:angleResult}.
 
@@
 	\includegraphics[width=2.5in]{images/angle_result}
-	\caption{An example result of the angle attribute map/}
+	\caption{An example result generated using the angle attribute map. Surfaces that are facing the camera have more detail, while those angled away from the camera are more transparent.}
 	\label{fig:angleResult}
@@
 We extend the shader to support the alpha channel of a tone detail texture to allow the use of transparency.
-Our method allows blending with original model's texture described in Section~\ref{sec:blending} or with the original method of using colour directly from the tone detail map.
-An example of transparency in a tone detail map can be seen in Figure~\ref{fig:transparency}.
+Our method allows blending with original model's texture described in Section~\ref{sec:blending} or with the original method of using colour directly from the tone detail texture.
+An example of transparency in a tone detail texture can be seen in Figure~\ref{fig:transparency}.
 This example uses the tone detail texture shown in Figure~\ref{fig:transparent_tone}.
@@
   \includegraphics[width=5.5in]{images/test}
-  \caption{Using a 2D texture from the original paper, we create a unique result by mixing the model's texture colors and the colors from the detail map.}
+  \caption{Using a 2D texture from the original paper, we create a unique result by mixing the model's texture colors and the colors from the tone detail texture.}
 \end{figure*}

### 2013-04-15 paper/paper.tex
Added new images and descriptions to paper Added new section to future work

--- 2013-04-15 paper/paper.tex
@@
 	\includegraphics[width=5.5in]{images/inverse_lighting}
-	\caption{Inverse Lighting result}
+	\caption{Inverse Lighting result. The model appears to be lit from the opposite direction but no lighting variables or directions have been altered.}
 	\label{fig:inverseLighting}
@@
 	\includegraphics[width=5.5in]{images/lighting}
-	\caption{Lighting result}
+	\caption{Lighting result. The model appears to self-shadow itself using only the tone detail texture.}
 	\label{fig:lightingResult}
@@
 	\includegraphics[width=1.5in]{images/xtoon_shading_gradient}
-	\caption{Tone texture for lighting}
+	\caption{Tone texture for lighting. This gives the effect of self-shadowing on a model, especially when combined with lighting directions.}
 	\label{fig:gradient}
@@
  \label{fig:alpha_issue}
+\end{figure*}
+
+\begin{figure*}[h]
+ \centering
+ \includegraphics[width=5.5in]{images/dissolve}
+ \caption{A dissolving effect. Using the lighting directions and the generated tone detail texture we were able to give the impression of the model dissolving in and out of view.}
+ \label{fig:dissolve}
+\end{figure*}
+
+\begin{figure*}[h]
+ \centering
+ \includegraphics[width=5.5in]{images/metallic}
+ \caption{A metallic sheen effect. The light directions give the model a metallic look as the tone detail texture reacts to new directions.}
+ \label{fig:metallic}
 \end{figure*}
@@
 
+\subsection{Animated tone detail texture}
+We also discussed the possibility of using an animated tone detail texture in lieu of a static image.
+This could create many new effects such as hologram noise and colour shifting without the need to change tone detail or shading values.
+Futures works should investigate the use of an animated tone detail texture and devise new effects that can be generated with this method.
+
 \section{Conclusion}

### 2013-04-15 paper/paper.tex
Small editing
Final editing

@@
 \usepackage{hyperref}
-\title{X\texorpdfstring{\textsuperscript{2}}-Toon: An Extended X-Toon Shader}
+\title{X\texorpdfstring{\textsuperscript{2}}.-Toon: An Extended X-Toon Shader}
 
@@
 
-The X-toon shader ~\cite{BTM06a} extended classic cel shading from 1D shading texture to 2D texture.
+The X-toon shader ~\cite{BTM06a} extended classic cel shading from 1D shading textures to 2D textures.
 The second dimension ``corresponds to the desired `tone detail' ''.
@@
  \centering
+ \includegraphics[width=5.5in]{images/alpha_issues}
+ \caption{Issues with alpha blending. The vertices on the front of the model can still be seen.}
+ \label{fig:alpha_issue}
+\end{figure*}
+
+\begin{figure*}[h]
+	\centering
 	\includegraphics[width=5.5in]{images/hologram}
@@
   \caption{Using a 2D texture from the original paper, we create a unique result by mixing the model's texture colors and the colors from the tone detail texture.}
-\end{figure*}
-
-\begin{figure*}[h]
- \centering
- \includegraphics[width=5.5in]{images/alpha_issues}
- \caption{Issues with alpha blending. The vertices on the front of the model can still be seen.}
- \label{fig:alpha_issue}
 \end{figure*}
@@
 This could create many new effects such as hologram noise and colour shifting without the need to change tone detail or shading values.
-Futures works should investigate the use of an animated tone detail texture and devise new effects that can be generated with this method.
+Future works should investigate the use of an animated tone detail texture and devise new effects that can be generated with this method.
 
